%%%%%%%%%%%%%%%%%%%%%%%%%% Packages / Definitions %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass{webofc}
\usepackage[varg]{txfonts}   % Web of Conferences font
\RequirePackage{orcidlink}
\usepackage{tikz}
\usetikzlibrary{calc, positioning, arrows.meta}%

\definecolor{problemred}{RGB}{140, 62, 84}   % muted red, print-safe

\tikzset{
  problemnote/.style={note, text=problemred},
  problemarrow/.style={drift, draw=problemred}
}

%%%%%%%%%%%%%%%%%%%%%%%%%% Document %%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\title{An IaC-Framework for the future of CERN’s Infrastructure and Configuration Management}

\author{\firstname{Christopher} \lastname{Barnes}\orcidlink{0009-0003-1963-7992}\inst{1}\fnsep\thanks{\email{christopher.barnes@cern.ch}} \and
        \firstname{Daniel} \lastname{Juárez}\orcidlink{0000-0000-0000-0000}\inst{1}\fnsep\thanks{\email{daniel.juarez@cern.ch}} \and
        \firstname{Luis} \lastname{Fernandez Alvarez}\orcidlink{0000-0000-0000-0000}\inst{1}\fnsep\thanks{\email{luis.fernandez.alvarez@cern.ch}} \and
        \firstname{Giacomo} \lastname{Tenagali}\orcidlink{0000-0000-0000-0000}\inst{1}\fnsep\thanks{\email{giacomo.tenagali@cern.ch}}
}

\institute{CERN, Esplanade des Particules 1, 1217 Meyrin, Switzerland}

\abstract{
CERN operates a large and distributed computing environment in which provisioning, configuration, and operational state are handled by different systems. Since 2012, the IT department has invested heavily in bridging these areas under the Agile Infrastructure project. Open-source projects such as OpenStack, Puppet, and Foreman have been integrated with in-house services to offer a cohesive view to its users.

Although mature, this ecosystem still faces difficulties for traceability, reproducibility, and automation, particularly when deploying new services, scaling existing ones, or recovering from failures. It also presents challenges to expose more advanced cloud resources, such as load balancers or security features, in a consistent and seamless way to the end users, in addition to the integration effort required.

Infrastructure-as-Code (IaC) approaches like Terraform and its open-source fork, OpenTofu, give a declarative and version-controlled model which could combine the current procedures and provide a more coherent workflow. These tools also benefit from a mature ecosystem, with off-the-shelf modules and components that can be readily reused to streamline and standardize these procedures.

We propose a framework that integrates with CERN’s existing infrastructure and configuration management landscape through declarative definitions, custom providers for in-house services, and CI/CD-driven validation pipelines. Through a set of case studies, we show how the proposed framework could reduce operational complexity, support disaster recovery scenarios, and improve automation, auditability, cross-environment replicability, offering a foundation for the evolution of infrastructure and configuration management at CERN.
}

\maketitle

\section{Introduction}
\label{intro}

CERN operates a large computing infrastructure. Since 2012, the IT department has developed a robust configuration management (CM) system with technologies like Puppet and OpenStack, alongside various in-house services, to manage its computing services~\cite{bellinfra2012}. These existing systems function efficiently for their specific purposes, but the overarching infrastructure management remains disjointed, leading to challenges in areas such as traceability, reproducibility, and maintainability.

Server management procedures exist, but maintaining service operation documentation to reflect rapid changes in infrastructure, especially in an environment with frequent personnel rotation, presents a significant challenge. This slows down effective recovery or reconfiguration in cases where responsible personnel are unavailable. Infrastructure-as-Code (IaC) tools, through their declarative nature, allow documenting the desired state of infrastructure, enhancing traceability and mitigating these documentation challenges. Furthermore, existing CM tools like Puppet only manage server configurations, not the underlying infrastructure (e.g., the servers themselves)~\cite{loope2011managing}, which can lead to operational difficulties when the underlying server infrastructure is degraded or becomes unavailable.

In recent years, IaC has emerged as a critical model for consolidating infrastructure development, deployment, and automation~\cite{morris2025infrastructure}. Terraform and its community-driven fork, OpenTofu, are prominent IaC solutions that offer substantial potential to unify infrastructure provisioning and configuration in a reproducible and scalable manner~\cite{holkuziev2025infrastructure}.

The core motivation of this work is/was? to design and implement an IaC-based framework, that integrates with CERN's existing infrastructure management ecosystem. This framework should provide a unified, automated, and traceable method for managing resources distributed across various cloud regions. A successful integration is expected to not only reduce operational complexity and maintenance effort but also significantly improve automation, enhance reproducibility, and ease use cases such as Business Continuity / Disaster Recovery (BC/DR) and cross-environment replication. This framework will serve as a foundational step towards a more unified and efficient infrastructure management strategy at CERN.
\section{Current Infrastructure Landscape}
\label{infra}

CERN operates its own cloud infrastructure based on OpenStack~\cite{dupuis2025,savchenko2024preparing}. Beyond OpenStack, several in-house services manage the machines state, trust, and network configuration~\cite{martelli2018}. Foreman serves as the central External Node Classification (ENC) system for Puppet, maintaining hostgroup assignments, environments, and metadata consumed during catalog compilation.

Production systems are typically provisioned through an internal tool that encapsulates the above mentioned in-house services, Foreman and OpenStack to create cloud resource. Once these are created a cloud-init script executes to run Puppet for bootstrapping and lifecycle management~\cite{rubavicius2018}. Once the Puppet agent performs its first catalog request and the host transitions into ongoing configuration management.

\subsection{Pain Points}
\label{sec:painpoints}

The primary friction does not originate from deficiencies within individual tools but from the interfaces between them. State is distributed across Foreman, OpenStack, alert management, state management, and CERN's network database, with parts duplicated or derived independently. As changes propagate through different mechanisms, inconsistencies can arise to the point where manual correction is required. Figure~\ref{fig:state-boundaries} illustrates the typical drift boundaries between infrastructure layers.

\begin{figure}[tbp]
  \label{fig:state-boundaries}
  \centering
  \begin{tikzpicture}[
      font=\small,
      box/.style={
          draw, rounded corners, align=center,
          minimum width=9.2cm, minimum height=1.05cm, inner sep=3pt
        },
      drift/.style={-Latex, thick},
      note/.style={align=left, font=\footnotesize},
      node distance=0.55cm
    ]

    % --- Layers ---
    \node[box] (svc) {Service / Application layer\\(\footnotesize team-specific procedures, overrides)};
    \node[box, below=of svc] (cm) {Configuration management layer\\(\footnotesize Puppet + Foreman)};
    \node[box, below=of cm] (cloud) {Virtual infrastructure layer\\(\footnotesize OpenStack resources, metadata, lifecycle)};
    \node[box, below=of cloud] (phys) {Physical infrastructure layer\\(\footnotesize hardware inventory, network registration, lifecycle records)};

    % --- Explicit boundary anchors (invisible points) ---
    \coordinate (b1) at ($(svc.south)!0.5!(cm.north)$);
    \coordinate (b2) at ($(cm.south)!0.5!(cloud.north)$);
    \coordinate (b3) at ($(cloud.south)!0.5!(phys.north)$);

    % --- Vertical arrows indicating boundaries ---
    \draw[drift] (svc.south) -- (cm.north);
    \draw[drift] (cm.south) -- (cloud.north);
    \draw[drift] (cloud.south) -- (phys.north);

    % --- Right-side boundary annotations (anchored to boundaries) ---
    \node[problemnote, right=5cm of b1] (d1)
    {boundary drift:\\manual steps,\\non-uniform idempotency};
    \draw[problemarrow] (d1.west) -- (b1);

    \node[problemnote, right=5cm of b2] (d2)
    {boundary drift:\\classification vs\\resource state};
    \draw[problemarrow] (d2.west) -- (b2);

    \node[problemnote, right=5cm of b3] (d3)
    {boundary drift:\\registration and\\lifecycle coordination};
    \draw[problemarrow] (d3.west) -- (b3);

    % --- Left-side annotation: multiple stores of state ---
    \node[note, left=0.6cm of cm.west] (sot)
    {multiple\\stores of\\state};
    \draw[drift] (sot.east) -- (cm.west);

  \end{tikzpicture}
  \caption{Layered view of the current infrastructure stack.}
\end{figure}


Key limitations include: overlapping state representations across systems with delayed convergence; asynchronous change propagation with weak end-to-end guarantees; manual steps remaining in provisioning for registration, hardware tracking, and exceptional workflows; and limited end-to-end observability with slow feedback loops. The \texttt{ai-bs} tool itself follows a fire-and-forget model, performing no pre-flight checks on permissions or quota and failing silently on errors.

The absence of a single declarative model spanning virtual resources, service-level configuration, and supporting registrations limits end-to-end automation and makes convergence dependent on cross-system coordination. These constraints motivate the exploration of a Terraform/OpenTofu-based framework to reduce boundary drift, improve idempotency across layers, and increase traceability of change.

\begin{thebibliography}{}

\bibitem{rubavicius2018}
R. Rubavicius, \textit{CENF-EHN1 Computing Cluster Manual}, Tech. Rep., CERN (2018)

\bibitem{martelli2018}
E. Martelli, T. Cass, \textit{CERN Data Centre Network Architecture: proposed evolution and implementation plan}, Tech. Rep., CERN (2018)

\bibitem{savchenko2024preparing}
M. Savchenko, L. Atzori, N. Papakyprianou, M. Piszczek, A. Wiebalck, EPJ Web Conf. \textbf{295}, 07038 (2024)

\bibitem{dupuis2025}
M. Dupuis, EPJ Web Conf. \textbf{337}, 01148 (2025)

\bibitem{morris2025infrastructure}
K. Morris, \textit{Infrastructure as Code: Designing and Delivering Dynamic Systems for the Cloud Age} (O'Reilly Media, 2025)

\bibitem{holkuziev2025infrastructure}
D. Holkuziev, Univ. Lib. Eng. Technol. \textbf{2}, 3 (2025)

\bibitem{loope2011managing}
J. Loope, \textit{Managing Infrastructure with Puppet: Configuration Management at Scale} (O'Reilly Media, 2011)

\bibitem{bellinfra2012}
P. Andrade, T. Bell, J. van Eldik, G. Mccance, B. Panzer-Steindel, M. Santos, S. Schwickerath, J. Phys.: Conf. Ser. \textbf{396}, 042002 (2012)
\end{thebibliography}

\end{document}
