\section{Introduction}
\label{intro}

CERN operates a large computing infrastructure. Since 2012, the IT department has developed a robust configuration management (CM) system with technologies like Puppet and OpenStack, alongside various in-house services, to manage its computing services~\cite{bellinfra2012}. These existing systems function efficiently for their specific purposes, but the overarching infrastructure management remains disjointed, leading to challenges in areas such as traceability, reproducibility, and maintainability.

Server management procedures exist, but maintaining service operation documentation to reflect rapid changes in infrastructure, especially in an environment with frequent personnel rotation, presents a significant challenge. This slows down effective recovery or reconfiguration in cases where responsible personnel are unavailable. Infrastructure-as-Code (IaC) tools, through their declarative nature, allow documenting the desired state of infrastructure, enhancing traceability and mitigating these documentation challenges. Furthermore, existing CM tools like Puppet only manage server configurations, not the underlying infrastructure (e.g., the servers themselves)~\cite{loope2011managing}, which can lead to operational difficulties when the underlying server infrastructure is degraded or becomes unavailable.

In recent years, IaC has emerged as a critical model for consolidating infrastructure development, deployment, and automation~\cite{morris2025infrastructure}. Terraform and its community-driven fork, OpenTofu, are prominent IaC solutions that offer substantial potential to unify infrastructure provisioning and configuration in a reproducible and scalable manner~\cite{holkuziev2025infrastructure}.

The core motivation of this work is/was? to design and implement an IaC-based framework, that integrates with CERN's existing infrastructure management ecosystem. This framework should provide a unified, automated, and traceable method for managing resources distributed across various cloud regions. A successful integration is expected to not only reduce operational complexity and maintenance effort but also significantly improve automation, enhance reproducibility, and ease use cases such as Business Continuity / Disaster Recovery (BC/DR) and cross-environment replication. This framework will serve as a foundational step towards a more unified and efficient infrastructure management strategy at CERN.
